Tensor denoising and completion based on ordinal observations

### 2020-02-05 ICML/submit/ordinalT.tex
update paper

\section{Introduction}
Multidimensional arrays, a.k.a.\ tensors, arise in a variety of applications including recommendation systems~\citep{adomavicius2011context}, social networks~\citep{nickel2011three}, genomics~\citep{wang2017three}, and neuroimaging~\citep{zhou2013tensor}. There is a growing need to develop dimension reduction method for analyzing these noisy, high-dimensional datasets. In one direction, \emph{tensor denoising} -- which aims to recover a signal tensor from its noisy entries -- has gained increased attention in both theory and applications~\citep{xia2019sup,xie2016multispectral}. In a related direction, \emph{tensor completion} examines the minimum number of entries needed for a consistent recovery~\citep{ghadermarzy2018learning,ghadermarzy2019near}. A low-rank structure is often imposed to the signal tensor, which efficiently reduces the intrinsic dimension in both problems.

In the line of these attempts, a number of low-rank tensor estimation methods have been proposed, revitalizing classical methods such as CANDECOMP/PARAFAC (CP) decomposition~\citep{hitchcock1927expression} and Tucker decomposition~\citep{tucker1966some}. These tensor methods have treated the entries as continuous-valued. In many cases, however, we encounter datasets of which the entries are qualitative. For example, the Netflix problem records the ratings of users to movies over weeks. Each data is a rating on a nominal scale \{{\it very like, like, neutral, dislike, very dislike}\}. Another example is in the signal processing, where the digits are frequently rounded or truncated so that only integer values are available. The qualitative observations take values in a limited set of categories, making the learning problem harder compared to the finer, continuous observations.

@@

\begin{algorithm}[t]
        \caption{Ordinal tensor decomposition and completion}\label{alg}
        \begin{algorithmic}[]
            \STATE{\bfseries Input:}  \text{ Ordinal data tensor }
            $\mathcal{Y}\in [L]^{d_1\times\cdots\times d_K}$, rank $\mr\in \mathbb{N}_{+}^{K}$, entry-wise bound $\alpha\in \mathbb{R_+}$.
            \STATE{\bfseries Output:} $(\hat\Theta,\hat{\mb}) =  \argmax_{(\Theta,\mb)\in \mathcal{P}\times\mathcal{B}}  \mathcal{L}_{\mathcal{Y},\Omega}(\Theta,\mb).$
            \STATE Initialize core tensor $\mathcal{C}^{(0)}$, factor matrices $ \{\mM_k^{(0)}\}$, and cut-off points $\mb^{(0)}$.
            \FOR{$t = 1,2,\cdots,$}
            \FOR{$k = 1,2,\cdots,K$}
            \STATE{Update $\mM^{(t+1)}_k$ while fixing other blocks:}
            \STATE $\mM_k^{(t+1)}\gets\argmax_{\mM_k\in\mathbb{R}^{d_k\times r_K}}\mathcal{L}_{\mathcal{Y},\Omega}(\mM_k)$,
            \newline
             s.t. $\mnormSize{}{\Theta^{(t+1)}}\leq \alpha$, where $\Theta^{(t+1)}$ is the parameter tensor based on the current block estimates. 
            \ENDFOR
            \STATE {Update $\mathcal{C}^{(t+1)}$ while fixing other blocks:}
              \newline$\mathcal{C}^{(t+1)} \gets \argmax_{\mathcal{C}\in \mathbb{R}^{r_1\times\cdots\times r_K}}\mathcal{L}_{\mathcal{Y},\Omega}(\mathcal{C})$, s.t. $\mnormSize{}{\Theta^{(t+1)}}\leq \alpha.$
               \STATE {Update $\Theta^{(t+1)}$ based on the current block estimates:} 
                   \newline
                   $\Theta^{(t+1)} \gets \mathcal{C}^{(t+1)}\times_1\mM_1^{(t+1)}\cdots\times_K\mM_K^{(t+1)}.$
            \STATE {Update $\mb^{(t+1)}$ while fixing $\Theta^{(t+1)}$:}
              \newline
              $\mb^{(t+1)} \gets \argmax_{\mb\in \mathbb{R}^{L-1}}\mathcal{L}_{\mathcal{Y},\Omega}\big(\Theta^{(t+1)},\mb\big).$
            \ENDFOR
            \STATE \textbf{return}
            $(\hat \Theta,\hat \mb)$
        \end{algorithmic}
 \end{algorithm}

Before concluding this section, we comment on two implementation details. First, the problem~\eqref{eq:estimator} is non-convex, so Algorithm 1 usually has no theoretical guarantee on global optimality. Nevertheless, as shown in Section~\ref{sec:denosing}, the desired rate holds not only for the global optimizer, but also for the local optimizer with $\logl(\hat \Theta)\geq \logl(\trueT)$. In practice, we find the convergence point $\hat \Theta$ is often satisfactory, in that the corresponding objective $\logl(\hat \Theta)$ is close to and actually slightly larger than the objective evaluated at the true parameter $\logl(\trueT)$. Figure~\ref{fig:stability} shows the trajectory of the objective function that is output in the default setting of Algorithm 1, with the input tensor generated from probit model~\eqref{eq:model} with $d_1=d_2=d_3=d$ and $r_1=r_2=r_3=r$. The dashed line is the objective value at the true parameter $\logl(\trueT)$. It is observed that the algorithm generally converges quickly to a desirable value in reasonable number of steps. The actual running time per iteration is shown in the plot legend.

@@

The first experiment examines the performance under complete observations. We assess the empirical relationship between the MSE and various aspects of model complexity, such as dimension $d$, rank $r$, and signal level $\alpha=\mnormSize{}{\Theta}$. Figure~\ref{fig:finite}a plots the estimation error versus the tensor dimension $d$ for three different ranks $r\in\{3,5,8\}$. The decay in the error appears to behave on the order of $d^{-2}$, which is consistent with our theoretical results~\eqref{eq:rate}. It is seen that a higher rank leads to a larger error, as reflected by the upward shift of the curve as $r$ increases. Indeed, a higher rank implies a higher intrinsic dimension, thus increasing the difficulty of the estimation. Figure~\ref{fig:finite}b shows the estimation error versus the signal level under $d=20$. Interestingly, a larger estimation error is observed when the signal is either too small or too large. The non-monotonic behavior may seem surprising, but this is in fact an intrinsic feature in the estimation with ordinal data. In view of the latent-variable interpretation (see Section~\ref{sec:latent}), estimation from ordinal observation can be interpreted as an inverse problem of quantization. Therefore, the estimation error will diverge when the noise $\tE$ is negligible, because there will be no way to distinguish two different signal tensors, e.g., $\Theta_1=\ma_1\otimes \ma_2\otimes \ma_3$ and $\Theta_2=\text{sign}(\ma_1)\otimes \text{sign}(\ma_2)\otimes \text{sign}(\ma_3)$, from the quantized observations. This phenomenon is clearly contrary to the classical continuous-valued tensor problem~\cite{davenport2014,sur2018modern}.

The second experiment investigates the incomplete observations. We consider $L$-level tensors with $d=20$, $\alpha=10$ and choose a subset of tensor entries via uniform sampling. Figure~\ref{fig:finite}c shows the estimation error of $\hat \Theta$ versus the fraction of observation $\rho=|\Omega|/d^K$. As expected, the error reduces with increased $\rho$ or decreased $r$. Figure~\ref{fig:finite}d evaluates the impact of ordinal levels $L$ to estimation accuracy, under the setting $\rho=0.5$. It is seen that the performance improves as $L$ grows, especially from binary observations ($L=2$) to multi-level ordinal observations ($L\geq 3$). The result showcases the benefit of multi-level observations provide compared to binary observations.

@@

\begin{itemize}[itemsep=0pt,topsep=-2pt,leftmargin=*]
\item Continuous tensor decomposition ({\bf Continuous-T})~\cite{filipovic2015tucker,oh2018scalable} is the classical Tucker approximation method. 
\item One-bit tensor completion ({\bf 1bit-T})~\cite{ghadermarzy2018learning} is a max-norm penalized tensor learning method based on partial binary observations. 
\item Ordinal matrix completion ({\bf Ordinal-M})~\cite{bhaskar2016probabilistic} is a rank-constrained matrix estimation method based on noisy, quantized observations. 
\vspace{-.4cm}
\end{itemize}

@@

Figure~\ref{fig:compare} compares the prediction accuracy under the setting $\alpha=10$, $d=20$, and $r=5$. The problem size we considered is comparable to~\citet{ghadermarzy2018learning}. It is seen that our method outperforms the others in both MAD and MCR. In particular, methods built on multi-level observations ({\bf Ordinal-T}, {\bf Ordinal-M}, {\bf 1bit-category-T}) exhibit stable MCR over $\rho$ and $L$, whereas the others two methods ({\bf Continuous-T}, {\bf 1bit-sign-T}) generally fail except for $L=2$ (Figures~\ref{fig:compare}a-b). This observation highlights the necessity of modeling multi-level probabilities in classification task. Interestingly, although both {\bf 1bit-category-T} and our method {\bf Ordinal-T} behave similarly for binary tensors ($L=2$), the improvement of our method is substantial as $L$ increases (Figures~\ref{fig:compare}a and~\ref{fig:compare}c). One possible reason is that our method incorporates the intrinsic ordering among the $L$ levels via proportional odds assumption~\eqref{eq:logodd}, whereas {\bf 1bit-category-T} ignores the ordinal structure and dependence among the induced binary entries. Figures~\ref{fig:compare}c-d assess the prediction accuracy with sample size. We see a clear advantage of our method ({\bf Ordinal-T}) over the matricization ({\bf Ordinal-M}) in both complete and non-complete observations. When the observation fraction is small, e.g., ${|\Omega|/ d^K}=0.4$, the tensor-based completion shows $\sim$ 30\% reduction in error compared to the matricization.

\begin{figure}[t]
\includegraphics[width=8.3cm]{panel2.pdf}
\vspace{-.4cm}
\caption{Performance comparison in MSE (a-b) and MAD (c-d). (a, b) Prediction errors versus the number of ordinal levels $L$ when $\rho=0.8$. (c, d) Prediction errors versus sample complexity $\rho=|\Omega|/d^K$ when $L=5$. }\label{fig:compare}
\vspace{-.4cm}
\end{figure}

We also compare the methods by their abilities in predicting the median score, $y_\omega^{\text{median}}=\min\{\ell\colon\mathbb{P}(y_\omega=\ell)\geq 0.5\}$. Note that, under the latent variable model~\eqref{eq:latent} and Assumption~\ref{ass:link}, the median score can be directly interpreted as the quantized $\Theta$ without noise; i.e.\ $y_\omega^{\text{median}}=\sum_\ell \mathds{1}_{\theta_\omega\in(b_{\ell-1},b_\ell]}$. Similar results demonstrate the outperformance of our method (see the Supplement and Figure S1).

@@

We compare the goodness-of-fit of various tensor methods on the HPC data. Table~\ref{table:CV} summarizes the prediction error via 5-fold stratified cross-validation averaged over 10 runs. It is clearly seen that our method outperform the others especially in MAD.

\begin{table}[http]
\resizebox{.48\textwidth}{!}{
\begin{tabular}{c|c|c}
\Xhline{2\arrayrulewidth}
\multicolumn{3}{c}{Human Connectome Project (HCP) dataset}\\
\Xhline{2\arrayrulewidth}
Method                    & MAD & MCR\\
\hline
Ordinal-T (ours) & 0.1607 (0.0005)& 0.1606 (0.0005)\\
 Continuous-T&0.2530 (0.0002)&0.1599 (0.0002)\\
 1bit-sign-T&0.3566 (0.0010)&0.1563 (0.0010)\\
 \Xhline{2\arrayrulewidth}
\multicolumn{3}{c}{InCarMusic dataset}\\
\Xhline{2\arrayrulewidth}
Method                    & MAD & MCR\\
\hline
Ordinal-T (ours) & 1.37 (0.039) &  0.59 (0.009)\\
Continuous-T & 2.39 (0.152)&0.94 (0.027)\\
1bit-sign-T&1.39 (0.003)& 0.81 (0.005)\\
\hline
\end{tabular}
}
\caption{Comparison of prediction error in the HPC and InCarMusic analyses. Standard errors are reported in parentheses.}\label{table:CV} 
\vspace{-.4cm}
\end{table}

\subsection{InCarMusic recommendation system}
In the second application, we apply ordinal tensor completion to a recommendation system {\it InCarMusic}. {\it InCarMusic} is a mobile application that offers music recommendation to passengers of cars based on contexts~\cite{baltrunas2011incarmusic}. The data tensor consists of partially observed ratings from scale 1 to 5 of 42 users to 139 songs on 26 contexts. Our goal is to perform tensor completion to impute the unobserved entries in the $42\times 139\times 26$ ordinal tensor and thereby we can offer context-specific music recommendation to users. Table~\ref{table:CV} shows the averaged prediction error via 5 fold cross validation. Again, our method achieves the best prediction error among the three. The data tensor has 98\% missing rate which makes the accurate classification challenging. Nevertheless, our method correctly classifies half of the unseen entries.

\section{Conclusion}
We have developed a low-rank tensor estimation method based on possibly incomplete, ordinal-valued observations. A sharp error bound is established, and we demonstrate the outperformance of our approach compared to other methods. The work unlocked several directions of future research. One interesting question would be the inference problem, i.e.. to assess the uncertainty of the obtained estimates and the imputation. Other directions include the trade-off between (non)convex optimization and statistical/computational efficiency. While convex relaxation is a popular approach for matrix-based problem, computing convex regularizer could be NP hard for tensors of order $K\geq 3$~\cite{hillar2013most, friedland2018nuclear}. The interplay between computational efficiency and statistical accuracy in general tensor problems warrants future research.

>>>>

\section{Introduction}
Multidimensional arrays, a.k.a.\ tensors, arise in a variety of applications including recommendation systems~\citep{adomavicius2011context}, social networks~\citep{nickel2011three}, genomics~\citep{wang2017three}, and neuroimaging~\citep{zhou2013tensor}. There is a growing need to develop dimension reduction method for analyzing these noisy, high-dimensional datasets. The problem of tensor denoising -- which aims to recover a signal tensor from its noisy entries -- has gained increased attention in both theory and applications~\citep{xia2019sup,xie2016multispectral}. A related problem, tensor completion, examines the minimum number of entries needed for a consistent recovery~\citep{ghadermarzy2018learning,ghadermarzy2019near}. Low-rankness is often imposed to the signal tensor, which efficiently reduces the intrinsic dimension in both problems.

A number of low-rank tensor estimation methods have been proposed~\citep{kolda2009tensor,acar2010scalable}, revitalizing classical methods such as CANDECOMP/PARAFAC (CP) decomposition~\citep{hitchcock1927expression} and Tucker decomposition~\citep{tucker1966some}. These tensor methods treat the entries as continuous-valued. In many cases, however, we encounter datasets of which the entries are qualitative. For example, the Netflix problem records the ratings of users to movies over weeks. Each data is a rating on a nominal scale \{{\it very like, like, neutral, dislike, very dislike}\}. Another example is in the signal processing, where the digits are frequently rounded or truncated so that only integer values are available. The qualitative observations take values in a limited set of categories, making the learning problem harder compared to the finer, continuous observations.

@@

\begin{algorithm}[t]
        \caption{Ordinal tensor decomposition}\label{alg}
        \begin{algorithmic}[]
            \STATE{\bfseries Input:}  \text{ Ordinal data tensor }
            $\mathcal{Y}\in [L]^{d_1\times\cdots\times d_K}$, rank $\mr\in \mathbb{N}_{+}^{K}$, entry-wise bound $\alpha\in \mathbb{R_+}$.
            \STATE{\bfseries Output:} $(\hat\Theta,\hat{\mb}) =  \argmax_{(\Theta,\mb)\in \mathcal{P}\times\mathcal{B}}  \mathcal{L}_{\mathcal{Y},\Omega}(\Theta,\mb).$
            \STATE Initialize core tensor $\mathcal{C}^{(0)}$, factor matrices $ \{\mM_k^{(0)}\}$, and cut-off points $\mb^{(0)}$.
            \FOR{$t = 1,2,\cdots,$}
            \FOR{$k = 1,2,\cdots,K$}
            \STATE{Update $\mM^{(t+1)}_k$ while fixing other blocks:}
            \STATE $\mM_k^{(t+1)}\gets\argmax_{\mM_k\in\mathbb{R}^{d_k\times r_K}}\mathcal{L}_{\mathcal{Y},\Omega}(\mM_k)$,
            \newline
             s.t. $\mnormSize{}{\Theta^{(t+1)}}\leq \alpha$, where $\Theta^{(t+1)}$ is the parameter tensor based on the current block estimates. 
            \ENDFOR
            \STATE {Update $\mathcal{C}^{(t+1)}$ while fixing other blocks:}
              \newline$\mathcal{C}^{(t+1)} \gets \argmax_{\mathcal{C}\in \mathbb{R}^{r_1\times\cdots\times r_K}}\mathcal{L}_{\mathcal{Y},\Omega}(\mathcal{C})$, s.t. $\mnormSize{}{\Theta^{(t+1)}}\leq \alpha.$
               \STATE {Update $\Theta^{(t+1)}$ based on the current block estimates:} 
                   \newline
                   $\Theta^{(t+1)} \gets \mathcal{C}^{(t+1)}\times_1\mM_1^{(t+1)}\cdots\times_K\mM_K^{(t+1)}.$
            \STATE {Update $\mb^{(t+1)}$ while fixing $\Theta^{(t+1)}$:}
              \newline
              $\mb^{(t+1)} \gets \argmax_{\mb\in \mathbb{R}^{L-1}}\mathcal{L}_{\mathcal{Y},\Omega}\big(\Theta^{(t+1)},\mb\big).$
            \ENDFOR
            \STATE \textbf{return}
            $(\hat \Theta,\hat \mb)$
        \end{algorithmic}
 \end{algorithm}

We comment on two implementation details before concluding this section. First, the problem~\eqref{eq:estimator} is non-convex, so Algorithm 1 usually has no theoretical guarantee on global optimality. Nevertheless, as shown in Section~\ref{sec:denosing}, the desired rate holds not only for the global optimizer, but also for the local optimizer with $\logl(\hat \Theta)\geq \logl(\trueT)$. In practice, we find the convergence point $\hat \Theta$ is often satisfactory, in that the corresponding objective $\logl(\hat \Theta)$ is close to and actually slightly larger than the objective evaluated at the true parameter $\logl(\trueT)$. Figure~\ref{fig:stability} shows the trajectory of the objective function that is output in the default setting of Algorithm 1, with the input tensor generated from probit model~\eqref{eq:model} with $d_1=d_2=d_3=d$ and $r_1=r_2=r_3=r$. The dashed line is the objective value at the true parameter $\logl(\trueT)$. It is observed that the algorithm generally converges quickly to a desirable value in reasonable number of steps. The actual running time per iteration is shown in the plot legend.

@@

The first experiment examines the performance under complete observations. We assess the empirical relationship between the MSE and various aspects of model complexity, such as dimension $d$, rank $r$, and signal level $\alpha=\mnormSize{}{\Theta}$. Figure~\ref{fig:finite}a plots the estimation error versus the tensor dimension $d$ for three different ranks $r\in\{3,5,8\}$. The decay in the error appears to behave on the order of $d^{-2}$, which is consistent with our theoretical results~\eqref{eq:rate}. It is seen that a higher rank leads to a larger error, as reflected by the upward shift of the curve as $r$ increases. Indeed, a higher rank implies a higher intrinsic dimension, thus increasing the difficulty of the estimation. Figure~\ref{fig:finite}b shows the estimation error versus the signal level under $d=20$. Interestingly, a larger estimation error is observed when the signal is either too small or too large. The non-monotonic behavior may seem surprising, but this is an intrinsic feature in the estimation with ordinal data. In view of the latent-variable interpretation (see Section~\ref{sec:latent}), estimation from ordinal observation can be interpreted as an inverse problem of quantization. Therefore, the estimation error diverges in the absence of noise $\tE$, because it is impossible to distinguish two different signal tensors, e.g., $\Theta_1=\ma_1\otimes \ma_2\otimes \ma_3$ and $\Theta_2=\text{sign}(\ma_1)\otimes \text{sign}(\ma_2)\otimes \text{sign}(\ma_3)$, from the quantized observations. This phenomenon is clearly contrary to the classical continuous-valued tensor problem~\cite{davenport2014,sur2018modern}.

The second experiment investigates the incomplete observations. We consider $L$-level tensors with $d=20$, $\alpha=10$ and choose a subset of tensor entries via uniform sampling. Figure~\ref{fig:finite}c shows the estimation error of $\hat \Theta$ versus the fraction of observation $\rho=|\Omega|/d^K$. As expected, the error reduces with increased $\rho$ or decreased $r$. Figure~\ref{fig:finite}d evaluates the impact of ordinal levels $L$ to estimation accuracy, under the setting $\rho=0.5$. An improved performance is observed as $L$ grows, especially from binary observations ($L=2$) to multi-level ordinal observations ($L\geq 3$). The result showcases the benefit of multi-level observations provide compared to binary observations.

@@

\begin{itemize}[itemsep=-.5pt,topsep=-.5pt,leftmargin=*,partopsep=-.5pt]
\item Continuous tensor decomposition ({\bf Continuous-T})~\cite{filipovic2015tucker,oh2018scalable} is the classical Tucker approximation method. 
\item One-bit tensor completion ({\bf 1bit-T})~\cite{ghadermarzy2018learning} is a max-norm penalized tensor learning method based on partial binary observations. 
\item Ordinal matrix completion ({\bf Ordinal-M})~\cite{bhaskar2016probabilistic} is a rank-constrained matrix estimation method based on noisy, quantized observations. 
\end{itemize}

@@

Figure~\ref{fig:compare} compares the prediction accuracy under the setting $\alpha=10$, $d=20$, and $r=5$. The problem size we considered is comparable to~\citet{ghadermarzy2018learning}. We find that our method outperforms the others in both MAD and MCR. In particular, methods built on multi-level observations ({\bf Ordinal-T}, {\bf Ordinal-M}, {\bf 1bit-category-T}) exhibit stable MCR over $\rho$ and $L$, whereas the others two methods ({\bf Continuous-T}, {\bf 1bit-sign-T}) generally fail except for $L=2$ (Figures~\ref{fig:compare}a-b). This observation highlights the necessity of modeling multi-level probabilities in classification task. Interestingly, although both {\bf 1bit-category-T} and our method {\bf Ordinal-T} behave similarly for binary tensors ($L=2$), the improvement of our method is substantial as $L$ increases (Figures~\ref{fig:compare}a and~\ref{fig:compare}c). One possible reason is that our method incorporates the intrinsic ordering among the $L$ levels via proportional odds assumption~\eqref{eq:logodd}, whereas {\bf 1bit-category-T} ignores the ordinal structure and dependence among the induced binary entries. Figures~\ref{fig:compare}c-d assess the prediction accuracy with sample size. We see a clear advantage of our method ({\bf Ordinal-T}) over the matricization ({\bf Ordinal-M}) in both complete and non-complete observations. When the observation fraction is small, e.g., ${|\Omega|/ d^K}=0.4$, the tensor-based completion shows $\sim$ 30\% reduction in error compared to the matricization.

\begin{figure}[t]
\includegraphics[width=8.3cm]{panel2.pdf}
\vspace{-.4cm}
\caption{Performance comparison in MSE (a-b) and MAD (c-d). (a, b) Prediction errors versus the number of ordinal levels $L$ when $\rho=0.8$. (c, d) Prediction errors versus sample complexity $\rho=|\Omega|/d^K$ when $L=5$. }\label{fig:compare}
\vspace{-.5cm}
\end{figure}

We also compare the methods by their abilities in predicting the median level, $y_\omega^{\text{median}}=\min\{\ell\colon\mathbb{P}(y_\omega=\ell)\geq 0.5\}$. Under the latent variable model~\eqref{eq:latent} and Assumption~\ref{ass:link}, the median level is the quantized $\Theta$ without noise; i.e.\ $y_\omega^{\text{median}}=\sum_\ell \mathds{1}_{\theta_\omega\in(b_{\ell-1},b_\ell]}$. Similar results demonstrate the outperformance of our method (see the Supplement).

@@

We compare the goodness-of-fit of various tensor methods on the HPC data. Table~\ref{table:CV} summarizes the prediction error via 5-fold stratified cross-validation averaged over 10 runs. Our method outperform the others, especially in MAD.

\begin{table}[http]
\resizebox{.48\textwidth}{!}{
\begin{tabular}{c|c|c}
\Xhline{2\arrayrulewidth}
\multicolumn{3}{c}{Human Connectome Project (HCP) dataset}\\
\Xhline{2\arrayrulewidth}
Method                    & MAD & MCR\\
\hline
Ordinal-T (ours) & 0.1607 (0.0005)& 0.1606 (0.0005)\\
 Continuous-T&0.2530 (0.0002)&0.1599 (0.0002)\\
 1bit-sign-T&0.3566 (0.0010)&0.1563 (0.0010)\\
 \Xhline{2\arrayrulewidth}
\multicolumn{3}{c}{InCarMusic dataset}\\
\Xhline{2\arrayrulewidth}
Method                    & MAD & MCR\\
\hline
Ordinal-T (ours) & 1.37 (0.039) &  0.59 (0.009)\\
Continuous-T & 2.39 (0.152)&0.94 (0.027)\\
1bit-sign-T&1.39 (0.003)& 0.81 (0.005)\\
\hline
\end{tabular}
}
\vspace{-.2cm}
\caption{Comparison of prediction error in the HPC and InCarMusic analyses. Standard errors are reported in parentheses.}\label{table:CV} 
\vspace{-.3cm}
\end{table}

\subsection{InCarMusic recommendation system}
We apply ordinal tensor completion to a recommendation system {\it InCarMusic}. {\it InCarMusic} is a mobile application that offers music recommendation to passengers of cars based on contexts~\cite{baltrunas2011incarmusic}. Our goal is to perform tensor completion to impute the unobserved entries in the $42\times 139\times 26$ ordinal tensor and thereby we can offer context-specific music recommendation to users. The data tensor consists of 2,884 observed entries. Table~\ref{table:CV} shows the averaged prediction error via 5-fold cross validation. The high missing rate makes the accurate classification challenging. Nevertheless, our method achieves the best performance among the three.

\section{Conclusion}
We have developed a low-rank tensor estimation method based on possibly incomplete, ordinal-valued observations. A sharp error bound is established, and we demonstrate the outperformance of our approach compared to other methods. The work unlocked several directions of future research. One interesting question would be the inference problem, i.e.. to assess the uncertainty of the obtained estimates and the imputation. Other directions include the trade-off between (non)convex optimization and statistical/computational efficiency. While convex relaxations are popular approach for matrix-based problem, they are often slow in practice~\cite{chen2019non}. The interplay between computational efficiency and statistical accuracy in general tensor problems warrants future research.
